The XAI Paradox: Systems that Perform Well for the Wrong Reasons

### 2018-08-28 Paper.tex
Added xcolor package for compilation. Corrected double abstract

\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
%PDF Info Is Required:
  \pdfinfo{
/Title (The XAI Paradox: Systems that Perform Well for the Wrong Reasons)
/Author (Cor Steging, Lambert Schomaker, Bart Verheij)}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{The XAI Paradox: Systems that Perform Well for the Wrong Reasons}
\author{Cor Steging, Lambert Schomaker, Bart Verheij\\
Department of Artificial Intelligence, Bernoulli Institute\\
University of Groningen\\
}
\maketitle
\begin{abstract}
Many of the successful modern machine learning approaches can be described as ``black box'' systems; these systems perform well, but are unable to explain the reasoning behind their decisions. The emerging sub-field of Explainable Artificial Intelligence (XAI) aims to create systems that are able to explain to their users why they made a particular decision. Using artificial datasets whose internal structure is known beforehand, this study shows that the reasoning of systems that perform well is not necessarily sound. Furthermore, when multiple combined conditions define a dataset, systems can perform well on the combined problem and not learn each of the individual conditions. Instead, a confounding structure within the data is learnt that still allows making the correct decisions. With regards to the goal of creating explainable systems, however, unsound rationales could create irrational explanations which would be problematic for the XAI movement.
Many of the successful modern machine learning approaches can be described as ``black box'' systems; these systems perform well, but are unable to explain the reasoning behind their decisions. The emerging sub-field of Explainable Artificial Intelligence (XAI) aims to create systems that are able to explain to their users why they made a particular decision. Using artificial datasets whose internal structure is known beforehand, this study shows that the reasoning of neural networks and decisions trees that perform well are not necessarily sound. Furthermore, when multiple conditions define a dataset, systems can perform well on the combined problem yet not learn each of the individual conditions that make up the combined problem. Instead, a confounding structure within the data is learnt that still allows making the correct decisions. With regards to the goal of creating explainable systems, however, unsound rationales could create irrational explanations which would be problematic for the XAI movement.
\end{abstract}

>>>>

\usepackage[table]{xcolor}

\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
%PDF Info Is Required:
  \pdfinfo{
/Title (The XAI Paradox: Systems that Perform Well for the Wrong Reasons)
/Author (Cor Steging, Lambert Schomaker, Bart Verheij)}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{The XAI Paradox: Systems that Perform Well for the Wrong Reasons}
\author{Cor Steging, Lambert Schomaker, Bart Verheij\\
Department of Artificial Intelligence, Bernoulli Institute\\
University of Groningen\\
}
\maketitle
\begin{abstract}
Many of the successful modern machine learning approaches can be described as ``black box'' systems; these systems perform well, but are unable to explain the reasoning behind their decisions. The emerging sub-field of Explainable Artificial Intelligence (XAI) aims to create systems that are able to explain to their users why they made a particular decision. Using artificial datasets whose internal structure is known beforehand, this study shows that the reasoning of systems that perform well is not necessarily sound. Furthermore, when multiple combined conditions define a dataset, systems can perform well on the combined problem and not learn each of the individual conditions. Instead, a confounding structure within the data is learnt that still allows making the correct decisions. With regards to the goal of creating explainable systems, however, unsound rationales could create irrational explanations which would be problematic for the XAI movement.
%Many of the successful modern machine learning approaches can be described as ``black box'' systems; these systems perform well, but are unable to explain the reasoning behind their decisions. The emerging sub-field of Explainable Artificial Intelligence (XAI) aims to create systems that are able to explain to their users why they made a particular decision. Using artificial datasets whose internal structure is known beforehand, this study shows that the reasoning of neural networks and decisions trees that perform well are not necessarily sound. Furthermore, when multiple conditions define a dataset, systems can perform well on the combined problem yet not learn each of the individual conditions that make up the combined problem. Instead, a confounding structure within the data is learnt that still allows making the correct decisions. With regards to the goal of creating explainable systems, however, unsound rationales could create irrational explanations which would be problematic for the XAI movement.
\end{abstract}
